## HW1

**Total points: 100**

```{r setup-hw1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", cache = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
```

1. (30pt)

\br(1) Load data... (3pt). Calculate... (5pt).

```{r}
df = read.csv("hw1/raindata.csv") %>% tibble()
df1 = df %>% 
  group_by(ym = floor_date(mdy(Date), unit = "month")) %>% 
  summarise_if(is.double, sum)
kable(df1)
```

\br(2) Calculate... (8pt).

```{r}
df2 = df1 %>% 
  group_by(month = str_sub(ym, 6, 7)) %>% 
  summarise_if(is.double, mean)
kable(select(df2, -ym))
```

\br(3) Calculate... (8pt). Plot... (4pt).

```{r}
df3 = df2 %>% 
  rowwise(month) %>% 
  mutate(regional_avg = mean(c_across(Site.A:Site.E))) %>% 
  ungroup()
kable(select(df3, month, regional_avg))
ggplot(df3, aes(ymd(ym), regional_avg)) +
  geom_point() +
  geom_line() +
  labs(x = "year-month", y = "regional-average monthly rainfall") +
  theme_test()
```

\br(4) Identify... (2pt).

```{r}
# dry seasons are winter and spring
df3 %>% 
  filter(regional_avg < 50) %>% 
  pull(ym) %>% 
  month()
# wet seasons are summer and fall
df3 %>% 
  filter(regional_avg > 50) %>% 
  pull(ym) %>% 
  month()
```


2. (30pt)

\br(1) Plot... (10pt). 2pt for selected stations. 5pt for the plot. 3pt for the x-axis label.

```{r}
df = read.csv("hw1/Myanmar_Monthly_Rainfall.csv") %>% tibble()
df_pick = df[sample(1:54, 5),] %>%  # randomly pick 5 stations
  pivot_longer(cols = -Site, names_to = "month", values_to = "avg") %>% 
  mutate(
    Site = str_c("S", Site),
    month = factor(month, levels = str_to_upper(month.abb)),
  )
ggplot(df_pick, aes(month, avg, color = Site, group = Site)) +
  geom_line() +
  labs(x = NULL, y = "average monthly rainfall") +
  theme_test() +
  theme(
    legend.direction = "horizontal",
    legend.position = c(.3, .9),
  )
```

\br(2) Plot... (5pt).

```{r}
ggplot(df_pick, aes(month, avg)) +
  geom_boxplot() +
  labs(x = NULL, y = "average monthly rainfall") +
  theme_test()
```

\br(3) Calculate... (10pt).

```{r}
df_pick %>% 
  group_by(month) %>% 
  summarise(mean = mean(avg), sd = sd(avg)) %>% 
  kable()
```

\br(4) Identify... (5pt). 3pt for the values of outliers. 1pt for the upper whisker and 1pt for the lower whisker.

```{r}
values = df_pick %>% filter(month == "JUL") %>% pull(avg)
values
smy = unname(summary(values))
IQR = smy[5] - smy[2]
values[values > smy[5] + 1.5 * IQR | values < smy[2] - 1.5 * IQR] # values of outliers
max(smy[2] - 1.5 * IQR, min(values)) # lower whisker
min(smy[5] + 1.5 * IQR, max(values)) # upper whisker
```

3. (20pt)

\br(1) Calculate... (10pt). Mean 5pt and sd 5 pt.

```{r}
df = readRDS("hw1/gosat_xco2_East_Asia.rds") %>% 
  mutate(
    # ym = format(as.Date(mdate), "%Y-%m"),
    ym = floor_date(as.Date(mdate), unit = "month"),
  ) %>% 
  group_by(ym) %>%
  summarise(
    ym = unique(ym),
    mean_co2 = mean(mr.co2), 
    sd_co2 = sd(mr.co2),
    .groups = "drop"
  )
kable(df)
```

\br(2) Plot... (10pt).

```{r}
df_48 = df %>% filter(month(ym) %in% c(4, 8))
plt_label = "June 2009 - Jan 2016 (Land & Sea)\nMonthly average and standard deviation of FTX XCO2"
ggplot(df, aes(ym, mean_co2)) +
  geom_point() +
  geom_point(data = df_48, aes(color = factor(month(ym)))) +
  geom_line() +
  geom_errorbar(aes(ymax = mean_co2 + qnorm(0.95) / 2 * sd_co2, 
                    ymin = mean_co2 - qnorm(0.95) / 2 * sd_co2),
                color = "#EE5555") +
  annotate(
    geom = "text", x = ymd(20140501), y = 382, size = 3.5,
    label = plt_label
  ) +
  labs(x = "year", y = "CO2(ppm)") +
  scale_x_date(
    date_breaks = "1 year",
    date_minor_breaks = "1 month",
    labels = date_format("%Y"),
    limits = ymd('2009-01-01','2016-01-01')
  ) +
  scale_color_manual(
    name = NULL, 
    breaks = c(4, 8), 
    labels = c("April", "August"), 
    values =c("#E61717", "#1421DE")
  ) +
  theme_test() +
  theme(
    legend.position = c(.42, .23),
    legend.justification = c("right", "top"),
  )
```

4. (20pt)

\br(1) Plot... (5pt).

```{r}
df = read.csv("hw1/Wind_Speed.csv", header = FALSE, col.names = "wind_speed")
df = df %>% mutate(site = "farm")
ggplot(df, aes(site, wind_speed)) +
  geom_boxplot() +
  labs(x = NULL, y = "10-min average wind speed") +
  theme_test()
```

\br(2) Plot... (5pt).

```{r}
ggplot(df, aes(wind_speed)) +
  geom_histogram() +
  labs(x = NULL, y = "10-min average wind speed") +
  theme_test()
```

\br(3) Explain... (10pt).

For example, skew to the right.

## Appendix: All code for HW1

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```