## HW2

**Total points: 160**

```{r setup-hw2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", cache = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
```

1. (5pt)

\br(1) (2pt)

Sample space is $\{\text{red cube}, \text{yellow cube}\}$.

\br(2) (3pt)

$$
\begin{aligned}
P(\text{yellow cube}|\text{not a blue cube})&=
\frac{P(\text{yellow cube }\cap\text{ not a blue cube})}{P(\text{not a blue cube})}\\
&=\frac{15/(22+15+13)}{(22+15)/(22+15+13)}\\
&=\frac{15}{37}
\end{aligned}
$$

2. (5pt)

\br(1) (1pt)

```{r}
cubes = rep(c("red", "yellow", "blue"), c(22, 15,13))

sample_space = cubes[which(cubes != "blue")]
# or equivalently,
# sample_space = cubes[which(!cubes %in% "blue")]
# sample_space = cubes[!cubes %in% "blue"]
```

\br(2) (2pt)

```{r}
sample(sample_space, size = 10, replace = TRUE)
```

\br(3) (2pt)

```{r}
sample(sample_space, size = 10, replace = FALSE)
# or equivalently,
# sample(sample_space, size = 10)
```

3. (42pt)

\br(1) Total: (37pt). 8pt for $f_X(x)$, 8pt for $f_Y(y)$, 13pt for $f_Z(z)$, 8pt for plot.

$$
\begin{aligned}
F_X(40)&=f_X(12)\cdot(40-5)/2=1\\
f_X(12)&=2/35\\
F_Y(55)&=f_Y(35)\cdot(55-15)/2=1\\
f_Y(35)&=1/20
\end{aligned}
$$

$$
\begin{aligned}
f_X(x)&=
\left\{ \begin{array}{ll}
\frac{2}{35}\frac{1}{12-5}(x-5) && 5\leq x<12\\
\frac{2}{35}-\frac{2}{35}\frac{1}{40-12}(x-12) && 12\leq x<40\\
\ 0 && \text{o.w.}
\end{array} \right.\\
&=
\left\{ \begin{array}{ll}
\frac{2}{245}x-\frac{2}{49} && 5\leq x<12\\
\frac{-1}{490}x+\frac{4}{49} && 12\leq x<40\\
\ 0 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

$$
\begin{aligned}
f_Y(y)&=
\left\{ \begin{array}{ll}
\frac{1}{20}\frac{1}{35-15}(y-15) && 15\leq y<35\\
\frac{1}{20}-\frac{1}{20}\frac{1}{55-35}(y-35) && 35\leq y<55\\
\ 0 && \text{o.w.}
\end{array} \right.\\
&=
\left\{ \begin{array}{ll}
\frac{1}{400}y-\frac{3}{80} && 15\leq y<35\\
\frac{-1}{400}y+\frac{11}{80} && 35\leq y<55\\
\ 0 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

$$
\begin{aligned}
f_Z(z)&=P(A)f_X(z)+P(B)f_Y(z)\\
&=\frac{1}{3}f_X(z)+\frac{2}{3}f_Y(z)\\
&=\left\{ \begin{array}{ll}
\frac{1}{3}\frac{2}{35}\frac{1}{12-5}(z-5) && 5\leq z<12\\
\frac{1}{3}\Big(\frac{2}{35}-\frac{2}{35}\frac{1}{40-12}(z-12)\Big) && 12\leq z<15\\
\frac{1}{3}\Big(\frac{2}{35}-\frac{2}{35}\frac{1}{40-12}(z-12)\Big)+\frac{2}{3}\Big(\frac{1}{20}\frac{1}{35-15}(z-15)\Big) && 15\leq z<35\\
\frac{1}{3}\Big(\frac{2}{35}-\frac{2}{35}\frac{1}{40-12}(z-12)\Big)+\frac{2}{3}\Big(\frac{1}{20}-\frac{1}{20}\frac{1}{55-35}(z-35)\Big) && 35\leq z<40\\
\frac{2}{3}\Big(\frac{1}{20}-\frac{1}{20}\frac{1}{55-35}(z-35)\Big) && 40\leq z<55\\
0 && \text{o.w.}
\end{array} \right.\\
&=\left\{ \begin{array}{ll}
\frac{2}{735}z-\frac{2}{147} && 5\leq z<12\\
\frac{-1}{1470}z+\frac{4}{147} && 12\leq z<15\\
\frac{29}{29400}z+\frac{13}{5880} && 15\leq z<35\\
\frac{-23}{9800}z+\frac{233}{1960} && 35\leq z<40\\
\frac{-1}{600}z+\frac{11}{120} && 40\leq z<55\\
0 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

```{r}
f_Z = function(z) {
  case_when(
    z >= 5 & z < 12 ~ 2 / 735 * z - 2 / 147,
    z >= 12 & z < 15 ~ -1 / 1470 * z + 4 / 147,
    z >= 15 & z < 35 ~ 29 / 29400 * z + 13 / 5880,
    z >= 35 & z < 40 ~ -23 / 9800 * z + 233 / 1960,
    z >= 40 & z < 55 ~ -1 / 600 * z + 11 / 120,
    TRUE ~ 0
  )
}
df = tibble(
  z = seq(0, 80, 0.1),
  f_z = f_Z(z),
  F_z = map_dbl(z, ~ integrate(f_Z, 0, .x)$value)
)
ggplot(df, aes(z, f_z)) +
  geom_line() +
  geom_vline(xintercept = c(12, 15, 35, 40), linetype = "dotted") +
  scale_x_continuous(breaks = c(5, 12, 15, 35, 40, 55), name = "z") +
  scale_y_continuous(limits = c(0, 0.1), name = latex2exp::TeX("$\\f_Z(z)$")) +
  theme_test()
```

\br(2) (5pt)

```{r}
# continous code in (1), or separate the c.d.f as follows
# df = df %>% 
#   mutate(F_z = map_dbl(z, ~ integrate(f_Z, 0, .x)$value))
ggplot(df, aes(z, F_z)) +
  geom_line() +
  geom_vline(xintercept = c(12, 15, 35, 40), linetype = "dotted") +
  scale_x_continuous(breaks = c(5, 12, 15, 35, 40, 55), name = "z") +
  scale_y_continuous(limits = c(0, 1.0001), name = latex2exp::TeX("$\\F_Z(z)$")) +
  theme_test()
```

4. (28pt)

\br(1) Total: (19pt). 1pt for $a$, 6pt for $f_X(x)$, 6pt for $E(X)$, 6pt for $Var(X)$.

$$
\begin{aligned}
F_X(10+2a)&=0.4\cdot2a/2=1,\quad a=2.5
\end{aligned}
$$

$$
\begin{aligned}
f_X(x)=
\left\{ \begin{array}{ll}
0.4/2.5\cdot(x-10)  & & 10\leq x< 12.5,\\
0.4 - 0.4/2.5\cdot(x-12.5) & & 12.5\leq x<15,\\
0 & & \text{o.w.}
\end{array} \right.
\end{aligned}
$$

$$
\begin{aligned}
E(X)
&=\int^{\infty}_{-\infty}xf_X(x)dx\\
&=\int^{12.5}_{10} x\cdot0.4/2.5\cdot(x-10)dx+\int^{15}_{12.5} x\cdot(0.4 - 0.4/2.5\cdot(x-12.5))dx\\
&=5.833333 + 6.666667\\
&=12.5
\end{aligned}
$$

```{r}
integrate(function(x) {x * 0.4/2.5 * (x - 10)}, 10, 12.5)$value + integrate(function(x) {x * (0.4 - 0.4 / 2.5 * (x - 12.5))}, 12.5, 15)$value
```

Or we can say that since the distribution is symmetric at $b$, the mean of the distribution is $b$.


$$
\begin{aligned}
Var(X)
&=E((X-E(X))^2)\quad \text{(法一)}\\
&=\int^{\infty}_{-\infty}(x-E(X))^2f_X(x)dx\\
&=\int^{12.5}_{10} (x-12.5)^2\cdot0.4/2.5\cdot(x-10)dx+\int^{15}_{12.5} (x-12.5)^2\cdot(0.4 - 0.4/2.5\cdot(x-12.5))dx\\
&=E(X^2) - E(X)^2\quad \text{(法二)}\\
&=\int^{12.5}_{10} x^2\cdot0.4/2.5\cdot(x-10)dx+\int^{15}_{12.5} x^2\cdot(0.4 - 0.4/2.5\cdot(x-12.5))dx-12.5^2\\
&=1.04
\end{aligned}
$$

```{r}
# 法一
integrate(function(x) {(x - 12.5)^2 * 0.4/2.5 *(x - 10)}, 10, 12.5)$value + integrate(function(x) {(x - 12.5)^2 * (0.4 - 0.4 / 2.5 * (x - 12.5))}, 12.5, 15)$value
# 法二
integrate(function(x) {x^2 * 0.4/2.5 * (x - 10)}, 10, 12.5)$value + integrate(function(x) {x^2 * (0.4 - 0.4 / 2.5 * (x - 12.5))}, 12.5, 15)$value - 12.5^2
```

\br(2) (3pt)

$$
\begin{aligned}
f_Y(y)=
\left\{ \begin{array}{ll}
f_X(y)/0.5 = (0.4 - 0.4/2.5\cdot(y-12.5))/0.5 && 12.5\leq y< 15,\\
0 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

\br(3) Total: (6pt). 3pt for mean, 3pt for variance.

$$
\begin{aligned}
E(Y)
&=\int^{\infty}_{-\infty}yf_Y(y)dy\\
&=\int^{15}_{12.5} y\cdot(0.4 - 0.4/2.5\cdot(y-12.5))/0.5dy\\
&=13.33
\end{aligned}
$$

```{r}
integrate(function(x) {x * (0.4 - 0.4 / 2.5 * (x - 12.5)) / 0.5}, 12.5, 15)$value
```


$$
\begin{aligned}
Var(Y)
&=E((Y-E(Y))^2)\quad \text{(法一)}\\
&=\int^{\infty}_{-\infty}(y-E(Y))^2f_Y(y)dy\\
&=\int^{15}_{12.5} (y-12.5)^2\cdot(0.4 - 0.4/2.5\cdot(y-12.5))/0.5\ dy\\
&=E(Y^2) - E(Y)^2\quad \text{(法二)}\\
&=\int^{15}_{12.5} y^2\cdot(0.4 - 0.4/2.5\cdot(y-12.5))/0.5\ dy-13.33333^2\\
&=0.35
\end{aligned}
$$

```{r}
# 法一
integrate(function(x) {(x - 13.3333)^2 * (0.4 - 0.4 / 2.5 * (x - 12.5)) / 0.5}, 12.5, 15)$value
# 法二
integrate(function(x) {x^2 * (0.4 - 0.4 / 2.5 * (x - 12.5)) / 0.5}, 12.5, 15)$value - 13.3333^2
```

5. (26pt)

\br(1) (8pt)

Let $X_1,X_2,\dots,X_n$ be random variables of n independent unit step.

$$
\begin{aligned}
&P(X_i=1)=p,\\
&P(X_i=-1)=1-p,\\
\end{aligned}
$$

for $i=1,2,\dots,n$.

Let $X$ be the position after $n$ steps.

$$\frac{X+n}{2}=\frac{\sum^n_{i=1} X_i+n}{2}=\sum^n_{i=1}\Big(\frac{X_i+1}{2}\Big)$$

Note that for $i=1,2,\dots,n$,

$$
\begin{aligned}
&P(\frac{X_i+1}{2}=1)=p,\\
&P(\frac{X_i+1}{2}=0)=1-p.
\end{aligned}
$$

That is, $\frac{X_i+1}{2}\sim\text{Bernoulli}(p)$, for $i=1,2,\dots,n$. Moreover, $\frac{X+n}{2}$ represents the number of successes that occur in the $n$ independent trials, and the following shows that it is said to be a binomial random variable with parameters $(n,p)$.

$$
\begin{aligned}
P(\frac{X+n}{2}=k)
&=P\Bigg(\sum^n_{i=1}\Big(\frac{X_i+1}{2}\Big)=k\Bigg)\\
&=P\Bigg(\Big(\frac{X_i+1}{2}\Big)=k_i,\ i=1,2,\dots,n\ \Big|\sum^n_{i=1}k_i=k, k_i\in(0,1),\ i=1,2,\dots,n\Bigg)\\
&={n\choose k}p^k(1-p)^{n-k},
\end{aligned}
$$

for $k=0,1,\dots,n$.

$\frac{X+n}{2}\sim\text{Binomial}(n,p)$

\br(2) Total: (5pt). 4pt for proof, 1pt for the rest.

Use the fact that $E(aX+b)=aE(X)+b$.

$$
\begin{aligned}
E(\frac{X+n}{2})&=\frac{1}{2}(E(X)+n)=np\\
E(X)&=2np-n
\end{aligned}
$$

\br(3) Total: (5pt). 4pt for proof, 1pt for the rest.

Use the fact that $Var(aX+b)=a^2Var(X)$.

$$
\begin{aligned}
Var(\frac{X+n}{2})&=(\frac{1}{2})^2(Var(X))=np(1-p)\\
Var(X)&=4np(1-p)
\end{aligned}
$$


The prove of the expectation and variance of a binomial random variable.

Let $X\sim\text{Binomial}(n,p)$

$$
\begin{aligned}
E[X^k] 
&= \sum_{i=0}^n i^k {n \choose i}p^i(1-p)^{n-i}\\
&= \sum_{i=1}^n i^k {n \choose i}p^i(1-p)^{n-i}\\
&= \sum_{i=1}^n i^k \frac{n}{i}{n-1 \choose i-1}p^i(1-p)^{n-i}\\
&= np\sum_{i=1}^n i^{k-1}{n-1 \choose i-1}p^{i-1}(1-p)^{n-i}\\
&= np\sum_{i=0}^{n-1} (i+1)^{k-1}{n-1 \choose i}p^i(1-p)^{n-i-1}\\
&= npE[(Y+1)^{k-1}]
\end{aligned}
$$

where $Y\sim Binom(n-1,p)$.

$$
\begin{aligned}
E[X]
&= npE[1]\\
&= np
\end{aligned}
$$


$$
\begin{aligned}
E[X^2]
&= npE[Y+1]\\
&= np((n-1)p+1)\\
Var(X)
&= np((n-1)p+1)-(np)^2\\
&= -np^2+np\\
&= np(1-p)
\end{aligned}
$$

\br(4) (3pt)

Previously we know that

$$
\begin{aligned}
P(\frac{X+n}{2}=k)
&={n\choose k}p^k(1-p)^{n-k},
\end{aligned}
$$

and then given $p=0.6$ and $n=20$ we can derive that

$$
\begin{aligned}
P(X>8)
&=P(\frac{X+20}{2}>(8+20)/2)\\
&=\sum^{20}_{k=15}{20\choose k}0.6^k(1-0.6)^{20-k}\\
&=0.1256\\
P(X< -8)
&=P(\frac{X+20}{2}<(-8+20)/2)\\
&=\sum^{5}_{k=0}{20\choose k}0.6^k(1-0.6)^{20-k}\\
&=0.0016\\
P(X>8)&+P(X< -8)=0.1272
\end{aligned}
$$

```{r}
sum(dbinom(c(0:5, 15:20), 20, 0.6))
```

\br(5) (5pt)

```{r}
df = tibble(
  simulation_set = factor(rep(1:5, each = 21)),
  n = rep(0:20, times = 5),
  move = map_dbl(n, ~ case_when(
    .x == 0 ~ 0,
    TRUE ~ sample(c(-1, 1), 1, prob = c(1 - 0.6, 0.6))
  )),
) %>% 
  group_by(simulation_set) %>% 
  mutate(position = cumsum(move))

ggplot(df, aes(n, position, color = simulation_set)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(name = "time", breaks = seq(0, 20, 1)) +
  scale_y_continuous(name = "distance", breaks = seq(-20, 20, 1)) +
  theme_test()
```

6. (14pt)

\br(1) (3pt)

WLOG, let $b=0$ and we can know what $a$ is.

$$
\begin{aligned}
1&=\int^1_{-1}a\sqrt{1-x^2}dx\\
&=a(\frac{x}{2}\sqrt{1-x^2}+\frac{1}{2}\sin^{-1}x)|^1_{-1}\\
&=\frac{a\pi}{2},\\
a&=\frac{2}{\pi}
\end{aligned}
$$

\br(2) (2pt)

$$
\begin{aligned}
f_Y(y)=
\left\{ \begin{array}{ll}
2\frac{2}{\pi}\sqrt{1-(x-b)^2} && b\leq x\leq b+1,\\
0 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

\br(3) (7pt) 1pt for the mean, 6pt for the variance.

Since $X$ is symmetric, we know that $E(X)=b$ trivially.

$$
\begin{aligned}
Var(X)
&=E((X-E(X))^2)\\
&=E((X-b)^2)\\
&=\int^{b+1}_{b-1}(x-b)^2a\sqrt{1-(x-b)^2}dx\\
\text{let }&y=x-b,\ dy=dx\\
&=\int^{1}_{-1}y^2a\sqrt{1-y^2}dy\\
&=\frac{a}{8}(\sin^{-1}(1)-\sin^{-1}(-1))\\
&=\frac{a\pi}{8}\\
&=\frac{2}{\pi}\frac{\pi}{8}
\end{aligned}
$$

\br(4) (2pt)

$$
\begin{aligned}
Z&=5+3(X-b)=3X+5-3b\\
E(Z)&=E(3X+5-3b)\\
&=3E(X)+5-3b\\
&=3b+5-3b\\
&=5\\
Var(Z)&=Var(3X+5-3b)\\
&=9Var(X)\\
&=\frac{9}{4}
\end{aligned}
$$

7. (40pt)

\br(1) Total: (32pt). Each of the r.v., we get 1pt for the highest density, 6 pt for the p.d.f., 6pt for the c.d.f, 3pt for the plot.

$$
\begin{aligned}
F_X(15)&=15f_X(5)/2=1\\
f_X(5)&=\frac{2}{15}\\
f_X(x)&=
\left\{ \begin{array}{ll}
\frac{2}{15}\frac{1}{5}x && 0\leq x<5,\\
\frac{2}{15}-\frac{2}{15}\frac{1}{10}(x-5)=\frac{1}{5}-\frac{x}{75} && 5\leq x<15,\\
0 && \text{o.w.}
\end{array} \right.\\
F_X(x)&=
\left\{ \begin{array}{ll}
\frac{1}{75}x^2 && 0\leq x<5,\\
F_X(5)+\int^x_5(\frac{2}{15}-\frac{2}{15}\frac{1}{10}(x-5))=-\frac{1}{2}+\frac{x}{5}-\frac{x^2}{150} && 5\leq x<15,\\
1 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$

```{r}
F_X = function(x) {
  case_when(
    x >= 0 & x < 5 ~ x^2 / 75,
    x >= 5 & x < 15 ~ -1 / 2 + x / 5 - x^2 / 150,
    x >= 15 ~ 1,
    TRUE ~ 0
  )
}
df = tibble(
  x = seq(0, 20, 0.01),
  F_x = F_X(x),
)
ggplot(df, aes(x, F_x)) +
  geom_line() +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20), name = "x") +
  scale_y_continuous(limits = c(0, 1.0001), name = latex2exp::TeX("$\\F_X(x)$")) +
  theme_test()
```



$$
\begin{aligned}
F_Y(15)&=10f_Y(10)/2=1\\
f_Y(10)&=\frac{1}{5}\\
f_Y(y)&=
\left\{ \begin{array}{ll}
\frac{1}{5}\frac{1}{5}(y-5)=\frac{1}{25}(y-5) && 5\leq y<10,\\
\frac{1}{5}-\frac{1}{5}\frac{1}{5}(y-10)=\frac{3}{5}-\frac{1}{25}y && 10\leq y<15,\\
0 && \text{o.w.}
\end{array} \right.\\
F_Y(y)&=
\left\{ \begin{array}{ll}
\frac{y^2}{50}-\frac{y}{5}+\frac{1}{2} && 5\leq y<10,\\
F_Y(10)+\int^y_{10}(\frac{3}{5}-\frac{1}{25}y)=\frac{-y^2}{50}+\frac{3y}{5}-\frac{7}{2} && 10\leq y<15,\\
1 && \text{o.w.}
\end{array} \right.
\end{aligned}
$$


```{r}
F_Y = function(y) {
  case_when(
    y >= 5 & y < 10 ~ y^2 / 50 - y / 5 + 1 / 2,
    y >= 10 & y < 15 ~ -y^2 / 50 + 3 * y / 5 - 7 / 2,
    y >= 15 ~ 1,
    TRUE ~ 0
  )
}
df = tibble(
  y = seq(0, 20, 0.01),
  F_y = F_Y(y),
)
ggplot(df, aes(y, F_y)) +
  geom_line() +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20), name = "y") +
  scale_y_continuous(limits = c(0, 1.0001), name = latex2exp::TeX("$\\F_Y(y)$")) +
  theme_test()
```

\br(2) (8pt)

$$
\begin{aligned}
P(X>Y)
&=\int P(X>y|y)f_Y(y)dy\\
&=\int^{15}_5\Big(\int^{15}_yf_X(x)dx\Big)f_Y(y)dy\\
&=\int^{10}_5\Big(\int^{15}_yf_X(x)dx\Big)f_Y(y)dy+\int^{15}_{10}\Big(\int^{15}_yf_X(x)dx\Big)f_Y(y)dy\\
&=\int^{10}_5\Big(\int^{15}_y\frac{1}{5}-\frac{x}{75}dx\Big)(\frac{1}{25}(y-5))dy+\int^{15}_{10}\Big(\int^{15}_y\frac{1}{5}-\frac{x}{75}dx\Big)(\frac{3}{5}-\frac{y}{25})dy\\
&=\int^{10}_5(\frac{3}{2}-\frac{y}{5}+\frac{y^2}{150})(-\frac{1}{5}+\frac{y}{25})dy+\int^{15}_{10}(\frac{3}{2}-\frac{y}{5}+\frac{y^2}{150})(\frac{3}{5}-\frac{y}{25})dy\\
&=0.19
\end{aligned}
$$

```{r}
integrate(function(y) {(3 / 2 - y / 5 + y^2 / 150) * (-1 / 5 + y / 25)}, 5, 10)$value + 
  integrate(function(y) {(3 / 2 - y / 5 + y^2 / 150) * (3 / 5 - y / 25)}, 10, 15)$value
```

## Appendix: All code for HW2

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

